---
title: "Notebook_5_create_reservoir_tile_maps"
output: html_notebook
---

## Creating tiled maps of reservoirs from the tabular data and the shape files
## Warning: May take a little time, depending on your computer's graphic capabilities due to large output file resolution
### Tomasz Janus
tomasz.k.janus@gmail.com

```{r, warning=FALSE, echo=FALSE}
library(sf)
library(readxl)
library(dplyr)
library(tmap)
library(tmaptools)
```
## 1. Import and Filter Data
```{r}
# Import shape file with reservoirs
res_shp_file_path = file.path("bin","heet_outputs_MIN_LOW_PRI","reservoirs_updated.shp")
res_shp_data = sf::st_read(res_shp_file_path)
# Read excel file with results
results_all <- readxl::read_excel(
  file.path("outputs", "reemission", "combined", "combined_outputs.xlsx"))

# Filter results to only contain MIN_LOW_PRI scenario
results_filtered <- results_all %>% dplyr::filter(Scenario == "MIN_LOW_PRIM")
# Elevation data
elev_data <- read.csv(file.path("config", "elev.csv"))
# Add elevation information to the shape file
res_shp_data <- res_shp_data %>%
  inner_join(elev_data, by = c("name" = "name"), suffix = c("_shp", "_results"))
```
```{r}
non_irr_res_names <- results_filtered %>% 
    dplyr::filter(type %in% c('hydroelectric', 'multipurpose')) %>% 
    dplyr::select('Name')
irr_res_names <- results_filtered %>% 
    dplyr::filter(type %in% c('irrigation')) %>% 
    dplyr::select('Name')
non_irr_res_names <- non_irr_res_names$Name
irr_res_names <- irr_res_names$Name
```
## 2. Combine shape file data with emissions
```{r}
# Combine shape data with re-emission results
res_shp_data <- res_shp_data %>%
  inner_join(results_filtered, by = c("name" = "Name"), suffix = c("_shp", "_results"))
res_shp_hp_multi <- res_shp_data %>% dplyr::filter(name %in% non_irr_res_names)
res_shp_irr <- res_shp_data %>% dplyr::filter(name %in% irr_res_names)

split_index <- nrow(res_shp_irr) / 2
res_shp_irr_1 <- res_shp_irr[1:split_index, ]
res_shp_irr_2 <- res_shp_irr[(split_index + 1):nrow(res_shp_irr), ]

save_maps <- TRUE
```

### 3. Plot the hydroelectric and the multipurpose reservoirs
```{r}
tmap::tmap_mode('plot')
m <- tmap::tm_shape(shp=res_shp_hp_multi) + 
    tmap::tm_polygons(col='res_area', style='cont', title = 'Reservoir\nArea, km2', alpha = 1.00) +
    tmap::tm_facets(
        by="name", free.scales = FALSE, free.coords = TRUE, 
        drop.NA.facets = FALSE,
        drop.empty.facets = FALSE,
        free.scales.symbol.size = 0.01, ncol = 10) +
    tm_style("col_blind") +
    tmap::tm_layout(main.title.position = "center",
              legend.outside = TRUE,
              legend.outside.size = 0.1,
              outer.margins = c(0.02,0.02,0.02,0.02),
              legend.outside.position = "right",
              main.title = "Hydroelectric and Multipurpose Reservoirs", 
              fontfamily = 'serif', legend.show = TRUE)
if (save_maps == TRUE) {
  tmap_save(
    m, file.path("figures","maps","reservoir_tiles_hp_multi.jpg"), 
    width=12, height=7, dpi = 2400)
  tmap_save(
    m, file.path("figures","maps","reservoir_tiles_hp_multi.png"), 
    width=12, height=7, dpi = 2400)
}
m
```

## 2. Plot irrigation reservoirs 1/2
```{r}
tmap::tmap_mode('plot')
m <- tmap::tm_shape(shp=res_shp_irr_1) + 
    tmap::tm_polygons(col='res_area', style='cont', title = 'Reservoir\nArea, km2', alpha = 1.00) +
    tmap::tm_facets(
        by="name", free.scales = FALSE, free.coords = TRUE, 
        drop.NA.facets = FALSE,
        drop.empty.facets = FALSE,
        free.scales.symbol.size = 0.01, ncol = 10) +
    tm_style("col_blind") +
    tmap::tm_layout(main.title.position = "center",
              legend.outside = TRUE,
              legend.outside.size = 0.1,
              outer.margins = c(0.02,0.02,0.02,0.02),
              legend.outside.position = "right",
              main.title = "Irrigation reservoirs (1/2)", fontfamily = 'serif', legend.show = TRUE)
if (save_maps == TRUE) {
  tmap_save(
    m, file.path("figures","maps","reservoir_tiles_irrig_1.jpg"), 
    width=12, height=7, dpi = 2400)
  tmap_save(
    m, file.path("figures","maps","reservoir_tiles_irrig_1.png"), 
    width=12, height=7, dpi = 2400)
}
m
```
## 3. Plot irrigation reservoirs 2/2
```{r}
tmap::tmap_mode('plot')
m <- tmap::tm_shape(shp=res_shp_irr_2) + 
    tmap::tm_polygons(col='res_area', style='cont', title = 'Reservoir\nArea, km2', alpha = 1.00) +
    tmap::tm_facets(
        by="name", free.scales = FALSE, free.coords = TRUE, 
        drop.NA.facets = FALSE,
        drop.empty.facets = FALSE,
        free.scales.symbol.size = 0.01, ncol = 10) +
    tm_style("col_blind") +
    tmap::tm_layout(main.title.position = "center",
              legend.outside = TRUE,
              legend.outside.size = 0.1,
              outer.margins = c(0.02,0.02,0.02,0.02),
              legend.outside.position = "right",
              main.title = "Irrigation reservoirs (2/2)", fontfamily = 'serif', legend.show = TRUE)
if (save_maps == TRUE) {
  tmap_save(
    m, file.path("figures","maps","reservoir_tiles_irrig_2.jpg"), 
    width=12, height=7, dpi = 2400)
  tmap_save(
    m, file.path("figures","maps","reservoir_tiles_irrig_2.png"), 
    width=12, height=7, dpi = 2400)
}
m
```